---
title: "Deloitte Monitor- Analytics chanllenge"
author: "Xiaoyan Zhou"
date: "12/07/2018"
output:
  html_document:
    number_sections: yes
    smart: no
    theme: paper
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
---

<br>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(stargazer)
library(ggplot2)
library(glmnet)
library(randomForest)
library(tidyr)
```

1. Random Forest forcasting 201610 sales volumn

Product A
```{r}
G1 <- read.csv("D26387251.csv")

```
Split to training set, validation set and test set( for prediction)

```{r}
data <- G1
sapply(data,class)
train <- data[data$sales_small.WeekKey <= 201550,-c(1,2,3,21) ]
valid <- data[201550< data$sales_small.WeekKey & data$sales_small.WeekKey<= 201608,-c(1,2,3,21) ]
test <- data[ 201608< data$sales_small.WeekKey,-c(1,2,3,24)]
train <- na.omit(train)
valid<- na.omit(valid)


test$price_change_nw <-0
#test$price_change_nw <-1 # use this when predicting the impact on providing 1 pound promotions





valid_y <-valid$sales_small.SalesVolume_nw
test_y <-test$sales_small.SalesVolume_nw
```

Train on the training set # manually tune ntree and nodesize to minimize the mse, can write for loop to test more combination
```{r}

set.seed(100)
rf <- randomForest(sales_small.SalesVolume_nw ~ ., train, ntree = 100,nodesize = 10)
imp <-importance(rf)
pred_value<- predict(rf, valid)
mse_rf <-mean((pred_value-valid_y)^2)
mse_rf

```

Feature importance in forcasting product A
```{r}
imp <-as.data.frame(imp)
imp$IncNodePurity <-round(imp$IncNodePurity, digit= 3)

area.color <- c("tomato1", "tomato1","tomato1","tomato1","tomato1", "tomato1", "tomato1", "skyblue2", "skyblue2","skyblue2","skyblue2",
                "skyblue2","skyblue2","skyblue2","skyblue2","skyblue2","skyblue2","skyblue2","skyblue2")
ggplot(aes(x = reorder(row.names(imp),-IncNodePurity), y = IncNodePurity), data =imp) +
  geom_bar(stat = 'identity', fill=area.color,width = 0.6) +
  theme_minimal()+
  labs(x='Features',  y = "Importance")+
  ggtitle("Feature importance in random forest algorithm - ProductA")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

next week prediction
```{r}
# train on training and validation set
train <- data[data$sales_small.WeekKey<= 201608,-c(1,2,3,24) ]
train <- na.omit(train)
set.seed(100)
rf <- randomForest(sales_small.SalesVolume_nw ~ ., train, ntree = 100,nodesize = 10)

pred_value<- predict(rf, test)
pred_value <- round(pred_value)
```



```{r}
historic <- data[data$sales_small.WeekKey <= 201609 &data$sales_small.WeekKey >= 201600,c(3,7) ]
colnames(historic) <-c("Week","D26387251")

predict_week <- rep(201610, length(pred_value))

prediction <- data.frame(predict_week, pred_value)
colnames(prediction)<- c("Week","D26387251")
```

```{r}
sales_volume_D26387251<- rbind(historic, prediction)
PA <-aggregate(sales_volume_D26387251$D26387251, by=list(Category=sales_volume_D26387251$Week), FUN=sum)
```


Change the product code and repeat the analysis for the other 3 products, write the prediction to a csv filr

```{r}
#sales_volume_Dbca94c97<- rbind(historic, prediction)
#PB <-aggregate(sales_volume_Dbca94c97$Dbca94c97, by=list(Category=sales_volume_Dbca94c97$Week), FUN=sum)

#sales_volume_Dedf80f3a<- rbind(historic, prediction)
#PC <-aggregate(sales_volume_Dedf80f3a$Dedf80f3a, by=list(Category=sales_volume_Dedf80f3a$Week), FUN=sum)


#sales_volume_D606565a1<- rbind(historic, prediction)
#PD <-aggregate(sales_volume_D606565a1$D606565a1, by=list(Category=sales_volume_D606565a1$Week), FUN=sum)

#sales_volume <-data.frame(Week = PA$Category, D26387251 = PA$x, Dbca94c97 = PB$x, Dedf80f3a = PC$x, D606565a1 = PD$x)
#sales_volume 
#write.csv(sales_volume, "sales_volume.csv")
#write.csv(sales_volume, "sales_volume_reduce_price.csv")
```

Add one pound to the price change nextweek and redo the steps before to predict the sales volume after promotion, write it to a csv file
```{r}
#test$price_change_nw <-1 # use this when predicting the impact on providing 1 pound promotions
#write.csv(sales_volume, "sales_volume_reduce_price.csv")

```



2. Graphs for forcasting

produce a dataframe that can use to make graph
```{r}
forcast <-read.csv("sales_volume.csv")
reduce_forcast <- read.csv("sales_volume_reduce_price.csv")

forcast$Week <- as.character(forcast$Week)
forcast$Week[forcast$Week== "201609"] <- "Week9"
forcast$Week[forcast$Week== "201610"] <- "Week10-Forcast"

reduce_forcast$Week <- as.character(reduce_forcast$Week)
reduce_forcast$Week[reduce_forcast$Week== "201610"] <- "Week10-Promote"


compare <-rbind(forcast[9:10,2:6],reduce_forcast[10,2:6])
compare
```

201609 and 201610 compare
```{r}
forcast_pure <-forcast[9:10,2:6]
compare_DF1 <- forcast_pure %>% gather(Product, Sales_volume, D26387251:D606565a1)
compare_DF1$Week <- factor(compare_DF1$Week, levels = c("Week9", "Week10-Forcast"))

area.color <- c("tomato1","skyblue2","#9999CC")

ggplot(compare_DF1, aes(fill=Week, y=Sales_volume, x=Product)) + 
    geom_bar(position="dodge", stat="identity")+ scale_fill_manual(values = area.color)+
  theme_minimal()+
  geom_text(aes(label=Sales_volume),position = position_dodge(width = 1), vjust=-0.8, color="black", size=3)+
  scale_x_discrete(labels=c("A","D","B","C" ))+
  labs(x='Products',  y = "Sales Volume")+
  ggtitle("201610 Sales Volume Forcast")
```


```{r}

compare_DF <- compare %>% gather(Product, Sales_volume, D26387251:D606565a1)
compare_DF$Week <- factor(compare_DF$Week, levels = c("Week9", "Week10-Forcast", "Week10-Promote"))
compare_DF 
```


201609 and 201610 and promotion compare
```{r}

area.color <- c("tomato1","skyblue2","#9999CC")

ggplot(compare_DF, aes(fill=Week, y=Sales_volume, x=Product)) + 
    geom_bar(position="dodge", stat="identity")+ scale_fill_manual(values = area.color)+
  theme_minimal()+
  geom_text(aes(label=Sales_volume),position = position_dodge(width = 1), vjust=-0.8, color="black", size=3)+
  scale_x_discrete(labels=c("A","D","B","C" ))+
  labs(x='Products',  y = "Sales Volume")+
  ggtitle("201610 Sales Volume Forcast - Promotion")
```


3. price elasticity 
```{r}
price_elasticity <- read.csv("product_elasticity.csv")

```


```{r}
means <- aggregate(elasticity ~  products.Group,price_elasticity, mean)
means$elasticity <- round(means$elasticity, 1)
ggplot(price_elasticity, aes(x=products.Group, y=elasticity, color=products.Group)) +geom_boxplot()+
  geom_text(data = means, aes(label = elasticity, y = elasticity  -2.2))+
  theme_minimal()+
  scale_x_discrete(labels=c("A","E","F","D","B","G","C" ))+
  labs(x='Products',  y = "Price Elasticity")+
  ggtitle("Price Elasticity")+
  theme(legend.position="none")
  
```




